---
title: "Kinematics Example"
author: "Nathan Cooper"
date: "November 16, 2017"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Kinematic Equations

+ Equations that allow us to track the position of an object.
+ Derived from the definitions of acceleration and velocity.
+ Assume a constant acceleration.

## Calcualting Velocity

Consider the defintion of velocity in one dimension.
$$
\frac{dv}{dt} = a \\
dv = a\space dt \\
\int dv = \int a\space dt \\
\int dv = a\int  dt \\
v = v_0 + a\space t
$$

Where a is the ojects acceleration, v is the speed at the end of the time interval t, $v_0$ is the initial velocity.

## Calculating the Position

Now consider the definition of velocity in 1 dimension.
$$
\frac{dx}{dt} = v \\
dx = v\space dt \\
\int dx = \int v\space dt \\
\int dx = \int (v_0 + a\space t) dt \\
\int dx = \int v_0 dt+ a\space \int t dt\\
x = x_0 +v_0t +\frac{1}{2}at^2
$$

x is the position of the object at the end of time interval t. $x_0$ is the initial position.

## In Practice

+ The above derivations were for 1 dimension.
+ 2 or 3 dimensional scenerios are just 2 or 3 1D equations.
+ Subscripts help keep track of different dimensions, e.g., $a_y$ or $v_z$.
+ Trigonometric functions may be needed to seperate dimensions.

## Example

A baseball is hit with initial velocity of 14.1 m/s at a 45 degree angle above the ground. How far from home plate does it travel before it is caught at the same height that it was hit. Treat homeplate as $x_0 = 0$m.

## First calculate y for flight time.
$$
v_{y,0} = 14.1*sin(45^o) \\
v_{y,0} = 9.97\space\frac{m}{s}\\
a = -9.8\space\frac{m}{s^2} \\
y = y_0 +v_{y,0}t +\frac{1}{2}at^2 \\
y - y_0  = 9.97t - 4.9t^2 \\
4.9t^2 = 9.97t \\
t = \frac{9.97\frac{m}{s}}{4.9\frac{m}{s^2}} \\
t = 2.03 s
$$

## Then calculate x for distance
$$
v_{x,0} = 14.1*sin(45^o) \\
v_{x,0} = 9.97\space\frac{m}{s}\\
a = 0\space\frac{m}{s^2} \\
x = x_0 +v_{x,0}t +\frac{1}{2}at^2 \\
x = 9.97\space\frac{m}{s} 2.03s  \\
x = 20.2 m
$$

## A graph of the system

```{r,eval=TRUE}
t = c(0:203)/100
x = 9.97*t
y = 9.97*t - 4.9*t^2
plot(x,y, xlab = "x (m)", ylab = "y (m)", col="orangered")
```